---
title: "Scedastic Surrogate Swarm R Package - Fitting Radial Basis Functions"
output: html_document
vignette: >
  %\VignetteIndexEntry{large_optimization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we will show how SSSO can be used to fit radial basis functions to some generated data. 

```{r setup}
rm(list=ls())
library(ScedasticSurrogateSwarmR)
set.seed(128)
```

## Problem Setup
The model we will be trying to fit will be a simple set of radial basis functions (RBF):
```{r}
model_function <- function(params){
  parammat=matrix(params,nrow=3)
  centers=parammat[1,]
  weights=parammat[2,]
  widths=parammat[3,]*10
  modout=rep(0,length(xdat))
  for (ix in 1:length(centers)){
    modout=modout+weights[ix]*exp(-(widths[ix]*abs(xdat-centers[ix]))^2)
  }

  return(modout)
}

#Equality Constraints
eq <- function(params){
  constraints = c(
  )
  return(constraints)
}

#Inequality Constraints
# of form <= 0
ineq <- function(params){
  constraints = c(
  )
  return(constraints)
}
```

We can set up our input values in the following manner:
```{r}
ndat=200
xdat=sort(runif(ndat,0,1))
params=c(0.15,0.8,0.5,0.8,0.8,0.5)
param_len = 6
ydat=model_function(params)
ydat[xdat<0.3]=ydat[xdat<0.3]+rnorm(sum(xdat<0.3),0,0.1)
xplot= c(xdat,seq(max(xdat),max(xdat)*1.1,length.out=0))

lowlim=rep(0,6)
highlim=rep(1.2,6)
```


## Swarm Optimization
Using a Scedastic Surrogate Swarm optimization:
```{r,animation.hook="gifski",fig.width=9,fig.height=5}
swarm_state <- initialize_swarm(desired_values = ydat,
                                param_len = 6,
                                lowlim=lowlim,
                                highlim=highlim,
                                ineq_w = c(),
                                eq_w = c(),
                                config = swarm.config(swarm_size = 30,
                                                      deg=2))

for (itt in 1:30){
  tempcontrol = swarm.control(poly_w=0.2,stoch_w = 0.2)
  swarm_state <- step_swarm(desired_values = ydat,
                            swarm_state,
                            ineq_w = c(),
                            eq_w=c(),
                            config = swarm.config(swarm_size=dim(swarm_state$x.p)[1],
                                                  num_cluster=1,
                                                  deg = 2,
                                                  swarm.control = tempcontrol))
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,swarm_state$pout,type="l",col="grey",xlab = NULL,ylab="value",ylim = c(0,1.2))
  points(xdat,ydat)
  surrogate_out <- surrogate_model(as.vector(swarm_state$poly_rec),swarm_state$polyouts,swarm_state$centersaves,deg=2)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(swarm_state$poly_rec)),col="red")
  title(paste0("Scedastic Surrogate Swarm Optimization, step = ",itt))
  legend(par('usr')[2], par('usr')[4], bty='n',
         xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}

print(paste("Number of function evals: ",dim(swarm_state$allxp)[1]))
```

## Search Optimization

We can also use the Scedastic Surrogate Swarm package to do non-swarm searching using surrogates:
```{r,animation.hook="gifski",fig.width=9,fig.height=5}
search_state <- initialize_search(param_len = 6,
                                  lowlim=lowlim,
                                  highlim=highlim,
                                  config = search.config(deg=1))

genstate = F
for (itt in 1:5){
  tempconfig = search.config(gen = genstate,
                             deg=1,
                             search_samples = 12,
                             search_mag=0.2,
                             revert_best = T)
  search_state <- step_search(desired_values = ydat,search_state,ineq_w = c(),eq_w=c(),config = tempconfig)

  modout = model_function(as.vector(search_state$poly_recs[1,]))
  surrogate_out <- surrogate_model(as.vector(search_state$poly_recs[1,]),search_state$polyouts,search_state$centersaves,deg=1)
  if (mean(abs((modout-surrogate_out)/modout))>0.5){
    genstate = T
  }else{
    genstate = F
  }
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,search_state$current_pout,type="l",col="grey",xlab = NULL,ylab="value",ylim = c(0,1.2))
  points(xdat,ydat)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(search_state$poly_recs[1,])),col="red")
  title(paste0("Surrogate Search Optimization, step = ",itt))
  legend(par('usr')[2], par('usr')[4], bty='n', xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}
print(paste("Number of function evals: ",dim(search_state$xpins)[1]))
```


## Search-Initialized Scedastic Surrogate Swarm Optimization
We can then convert these search states and the learned surrogates into the initialization of a Scedastic Surrogate Swarm

```{r,animation.hook="gifski",fig.width=9,fig.height=5}
usevec = sample(1:dim(search_state$xpins)[1],24)

#Swap into particle swarm
swarm_state <- convert_search_to_swarm(search_state = search_state,
                                       usevec = usevec,
                                       desired_values = ydat,
                                       param_len = 6,
                                       lowlim=lowlim,
                                       highlim=highlim,
                                       ineq_w = c(),
                                       eq_w=c(),
                                       config = swarm.config(swarm_size=length(usevec),
                                                             num_cluster=1,
                                                             deg=1))

for (itt in 1:20){
  tempcontrol = swarm.control(poly_w=0.2,stoch_w = 0.2)
  swarm_state <- step_swarm(desired_values = ydat,
                            swarm_state,
                            ineq_w = c(),
                            eq_w=c(),
                            config = swarm.config(swarm_size=length(usevec),
                                                 num_cluster=1,
                                                 deg=1,
                                                 swarm.control = tempcontrol))
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,swarm_state$pout,type="l",col="grey",xlab = NULL,ylab="value",ylim = c(0,1.2))
  points(xdat,ydat)
  surrogate_out <- surrogate_model(as.vector(swarm_state$poly_rec),swarm_state$polyouts,swarm_state$centersaves,deg=1)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(swarm_state$poly_rec)),col="red")
  title(paste0("Search-Initialized Scedastic Surrogate Swarm Optimization, step = ",itt))
  legend(par('usr')[2], par('usr')[4], bty='n',
         xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}
print(paste("Number of function evals: ",dim(swarm_state$allxp)[1]))
```
