---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Scedastic Surrogate Swarm R Package Introduction

```{r setup}
rm(list=ls())
library(ScedasticSurrogateSwarmR)
set.seed(128)
```


## Problem Setup
The first step of using Scedastic Surrogate Swarm Optimization is to establish a model function we wish to optimize. 

```{r}
model_function <-function(params){
  modval<-params[1]*cos(xdat*params[2])+params[2]*sin(xdat*params[1])
  return(modval)
}
```

This package also expects us to specify inequality and equality constraints, but we can leave these blank. 

```{r}
#Equality Constraints
eq <- function(params){
  constraints = c(
  )
  return(constraints)
}

#Inequality Constraints
# of form <= 0
ineq <- function(params){
  constraints = c(
  )
  return(constraints)
}
```

Now let us establish some observe data we want to fit our model to:
```{r}
xdat=seq(1,100,0.5)
params=c(0.32,0.4)
ydat=modf(params)

lowlim=rep(0,2)
highlim=rep(1,2)
```


## Swarm Optimization
We can randomly initialize the position of particles in our swarm:
```{r}
swarm_state <- initialize_swarm(desired_values = ydat,param_len = 2,lowlim=lowlim,highlim=highlim,ineq_w = c(),eq_w=c(),config = swarm.config(swarm_size=100,num_cluster=3,deg=2))
```

Then we simply step through the optimization, graphing the progress as we wish:
```{r,animation.hook="gifski",fig.width=10,fig.height=5}
for (itt in 1:10){
  tempcontrol = swarm.control(poly_w=0.2,stoch_w = 0.01)
  swarm_state <- step_swarm(desired_values = ydat,swarm_state,ineq_w = c(),eq_w=c(),config = swarm.config(swarm_size=100,num_cluster=3,swarm.control = tempcontrol))
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,swarm_state$pout,type="l",col="grey",xlab = NULL,ylab="value")
  points(xdat,ydat)
  surrogate_out <- surrogate_model(as.vector(swarm_state$poly_rec),swarm_state$polyouts,swarm_state$centersaves,deg=2)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(swarm_state$poly_rec)),col="red")
  title("Scedastic Surrogate Swarm Optimization")
  legend(par('usr')[2], par('usr')[4], bty='n', xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}
print(paste("Number of function evals: ",dim(swarm_state$allxp)[1]))
```


## Search Optimization

We can also use the Scedastic Surrogate Swarm package to do non-swarm searching using surrogates:
```{r}
search_state <- initialize_search(param_len = 2,lowlim=lowlim,highlim=highlim,config = search.config())

genstate = F
for (itt in 1:10){
  tempconfig = search.config(gen = genstate,revert_best = T)
  search_state <- step_search(desired_values = ydat,search_state,ineq_w = c(),eq_w=c(),config = tempconfig)

  modout = model_function(as.vector(search_state$poly_recs[1,]))
  surrogate_out <- surrogate_model(as.vector(search_state$poly_recs[1,]),search_state$polyouts,search_state$centersaves,deg=2)
  if (mean(abs((modout-surrogate_out)/modout))>2){
    genstate = T
  }else{
    genstate = F
  }
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,search_state$current_pout,type="l",col="grey",xlab = NULL,ylab="value")
  points(xdat,ydat)
  surrogate_out <- surrogate_model(as.vector(search_state$poly_recs[1,]),swarm_state$polyouts,swarm_state$centersaves,deg=2)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(search_state$poly_recs[1,])),col="red")
  title("Surrogate Search Optimization")
  legend(par('usr')[2], par('usr')[4], bty='n', xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}
print(paste("Number of function evals: ",dim(search_state$xpins)[1]))
```


## Search-Initialized Scedastic Surrogate Swarm Optimization
We can then convert these search states and the learned surrogates into the initialization of a Scedastic Surrogate Swarm

```{r}
usevec = (dim(search_state$xpins)[1]-50):dim(search_state$xpins)[1]

#Swap into particle swarm
swarm_state <- convert_search_to_swarm(search_state = search_state,usevec = usevec,desired_values = ydat,param_len = 2,lowlim=lowlim,highlim=highlim,ineq_w = c(),eq_w=c(),config = swarm.config(swarm_size=length(usevec),num_cluster=3,deg=2))

for (itt in 1:10){
  tempcontrol = swarm.control(poly_w=0.2,stoch_w = 0.01)
  swarm_state <- step_swarm(desired_values = ydat,swarm_state,ineq_w = c(),eq_w=c(),config = swarm.config(swarm_size=length(usevec),num_cluster=3,swarm.control = tempcontrol))
  par(oma=c(0, 0, 0, 5))
  matplot(xdat,swarm_state$pout,type="l",col="grey",xlab = NULL,ylab="value")
  points(xdat,ydat)
  surrogate_out <- surrogate_model(as.vector(swarm_state$poly_rec),swarm_state$polyouts,swarm_state$centersaves,deg=2)
  lines(xdat,surrogate_out,col="blue")
  lines(xdat,model_function(as.vector(swarm_state$poly_rec)),col="red")
  title("Search-Initialized Scedastic Surrogate Swarm Optimization")
  legend(par('usr')[2], par('usr')[4], bty='n',
         xpd=NA,c("Observed","Modeled","Predicted","Actual"),fill=c("black","grey","blue","red"))
}
print(paste("Number of function evals: ",dim(swarm_state$allxp)[1]))
```

